---
title: "Untitled"
author: "Zachary M. Smith"
date: "January 3, 2018"
output: html_document
---


```{r echo=FALSE}
knitr::opts_chunk$set(eval=evaluate, cache=cache.me)
```

The BIBI data frame (`bibi.df`) is subset to only represent unique bioregion categories. This data frame contains all of the bioregions defined in the Chesapeake Bay basin by Smith et al (2017) __ADD REF__.
```{r}
bioregion.df <- bibi.df %>% 
  select(subspatial) %>% 
  filter(!subspatial %in% c("inland", "coast")) %>% 
  rename(bioregion =  subspatial) %>% 
  distinct()
```

Nest the data frame by sampling period (`period`), spatial resolution (`spatial`), and bioregion (`biregion`) in preparation of performing the bootstrap analyses (see [Nested Data Frame]).
```{r}
bibi.nest.bioregion <- nest_df(bibi.sub, bioregion.df, bioregion) 
```

Perform the bootstrap analyses for each sampling period (`period`), spatial resolution (`spatial`), and bioregion (`bioregion`) (see [Multiple Bootstrap Ratings]).
```{r}
boot.bioregion <- multi_bootstrap_ratings(bibi.nest.bioregion, boot.reps, boot.samp.sizes,
                                     spatial.vec, periods.vec)
```

Generate bar plots with standard deviation bars for each rating category by sampling period (`period`) and spatial resolution (`spatial`) (see [Bootstrap Bar Plots]).
```{r, fig.width = 8, fig.height = 7}
plot_boot_bar(boot.bioregion, periods.vec, title = "Bioregion")
```

Generate line plots of standard deviation bars for each rating category by bootstrap sampling size (see [Bootstrap Standard Deviation Plots]). Each plot represents a unique sampling period (`period`) and spatial resolution (`spatial`).
```{r, fig.width = 8, fig.height = 7}
plot_boot_sd(boot.bioregion, periods.vec, title = "Bioregion")
```

Generate area plots of the proportion of each rating after excluding samples classified as insufficient (see [Bootstrap Area Plots]). Each plot represents a unique sampling period (`period`) and spatial resolution (`spatial`).
```{r, fig.width = 15, fig.height = 7}
exclude_insufficient(boot.bioregion) %>% 
rating_area_plot("Bioregion")
```