---
title: "R Notebook"
output: html_document
---


```{r echo=FALSE}
knitr::opts_chunk$set(eval=run.cedr.acquisition)
```

```{r}
huc12.df <- data.table::fread("H:/Projects/Chessie_BIBI/report/FINAL_May25_2017/2017_Data/HUC12/TAB_HUC_12.csv") %>% 
  clean_up()
```

Export the HUC 12 data (`huc12.df`) as a CSV file to the "data" folder in this project's directory. Using a combination of `dir.create()`, `file.path()`, and `rprojroot::find_rstudio_root_file()` the necessary folder structure in the project directory will be created, if it does not already exist.
```{r}
dir.create(file.path(rprojroot::find_rstudio_root_file(), "data/huc12"),
           recursive = TRUE, showWarnings = FALSE)

data.table::fwrite(huc12.df, file.path(rprojroot::find_rstudio_root_file(), "data/huc12", "huc12.csv"))
```


```{r}
file.dir <- "H:/Projects/Chessie_BIBI/report/FINAL_May25_2017/2017_Data/Scores_Ratings/BIBI_Scores_Ratings_06292017.xlsx"
```

```{r}
bioregion.df <- read_excel(file.dir,
                           sheet = "Bioregion_Family",
                           col_types = c(
                             rep("text", 2),
                             "numeric",
                             rep("text", 2),
                             "date", 
                             "text",
                             rep("numeric", 2),
                             rep("text", 5),
                             rep("numeric", 5),
                             "text"
                           ),
                           na = c("", "NA", "na", " ")) %>% 
  clean_up()

region.df <- read_excel(file.dir,
                        sheet = "Region_Family",
                        col_types = c(
                          rep("text", 2),
                          "numeric",
                          rep("text", 2),
                          "date", 
                          "text",
                          rep("numeric", 2),
                          rep("text", 5),
                          rep("numeric", 5),
                          "text"
                        ),
                        na = c("", "NA", "na", " ")) %>% 
  clean_up()
```

```{r}
bibi.df <- bind_rows(region.df, bioregion.df) %>% 
  rename(subspatial = spatial) %>% 
  mutate(spatial = if_else(subspatial %in% c("coast", "inland"), 
                           "region", "bioregion")) %>% 
  select(spatial, everything())
```

Export the Chessie BIBI scores and ratings (`bibi.df`) as a CSV file to the "data" folder in this project's directory. Using a combination of `dir.create()`, `file.path()`, and `rprojroot::find_rstudio_root_file()` the necessary folder structure in the project directory will be created, if it does not already exist.
```{r}
dir.create(file.path(rprojroot::find_rstudio_root_file(), "data/raw"),
           recursive = TRUE, showWarnings = FALSE)

data.table::fwrite(bibi.df, file.path(rprojroot::find_rstudio_root_file(), "data/raw", "scores_ratings.csv"))
```

```{r}
rm(huc12.df, file.dir, bioregion.df, region.df, bibi.df)
```

