---
title: "Untitled"
author: "Zachary M. Smith"
date: "March 9, 2018"
output: html_document
---

```{r echo=FALSE}
knitr::opts_chunk$set(eval=evaluate, cache=cache.me)
```

## Assign Periods

Assign sampling periods of interest using `assign_periods()` ([Assign Sampling Periods]). This example, is centered aroun the potential 2008 baseline period represented by 2004-2008, a five year period. The remaining sampling periods are five year periods that can be formed with the current temporal range of the Chessie BIBI database.
```{r}
periods.vec <- c("1994_1998", "1999_2003", "2004_2008", "2009_2013")
bibi.periods <- assign_periods(bibi.df, periods.vec)
```

## Stacked Spatial Function

```{r}
spatial_stacked <- function(x) {
  x %>% 
    mutate(bioregion = str_replace(huc12, "_.*", "")) %>% 
  group_by(spatial_period, bioregion, rating, total) %>% 
  summarize(stream_miles = sum(stream_miles)) %>% 
  group_by(spatial_period) %>% 
  filter(rating != "Insufficient") %>% 
  mutate(total = sum(stream_miles)) %>% 
  group_by(spatial_period, bioregion) %>% 
  mutate(bioregion_total = sum(stream_miles),
         bioregion_percentage = stream_miles / bioregion_total * 100,
         basin_percentage = stream_miles / total * 100) %>% 
  ungroup() %>% 
    group_by(spatial_period, bioregion) %>% 
  summarize(percent = sum(basin_percentage)) %>% 
  ungroup() %>% 
  separate(spatial_period, c("spatial", "period1", "period2"), remove = FALSE) %>%
  unite(period, period1, period2) %>% 
  mutate(period = factor(period, levels = rev(periods.vec))) %>% 
  arrange(period) %>% 
  filter(spatial == "bioregion") %>% 
  ggplot(aes(period, percent, fill = bioregion)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = viridis::viridis_pal(option = "D")(12)) +
  coord_flip() +
    ylab("Percentage (%)") +
    xlab("Period")
}
```



## 1) No Spatial Aggregation

```{r, fig.width = 8, fig.height=5}
rating.no.spatial <- no_spatial_rating(bibi.periods, su.df) %>% 
  arrange(spatial_period) %>% 
  mutate(spatial_period = factor(spatial_period, levels = rev(unique(spatial_period)))) %>% 
  arrange(spatial_period)

no.spatial.bar <- plot_rating(rating.no.spatial, "Individual Sample Ratings", group.col = NULL)

no.spatial.bar
```

```{r, fig.width = 8, fig.height=10}

subspatial.df <- bibi.periods %>% 
  group_by(spatial_period, spatial, subspatial) %>% 
  summarize(count = n()) %>% 
  group_by(spatial_period) %>% 
  mutate(total = sum(count),
         percent = count / total * 100)

bio.plot <- subspatial.df %>% 
  filter(spatial == "bioregion") %>% 
  ggplot(aes(spatial_period, percent, fill = subspatial)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = viridis::viridis_pal(option = "D")(12)) +
  coord_flip() +
  facet_wrap(~spatial,
             scales = "free",
             nrow = 2)

region.plot <- subspatial.df %>% 
  filter(spatial == "region") %>% 
  ggplot(aes(spatial_period, percent, fill = subspatial)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = viridis::viridis_pal(option = "D")(2)) +
  coord_flip() +
  facet_wrap(~spatial,
             scales = "free",
             nrow = 2)

cowplot::plot_grid(bio.plot, region.plot, ncol = 1)
```
## 2) Watershed Mean Score


```{r}
mean.rating.df <- loop_cell_rating(bibi.periods, su.df,
                                   unit.col = "stream_miles",
                                   su.vec = c("huc12"),
                                   job = "mean_rating") %>% 
  arrange(spatial_period) %>% 
  mutate(spatial_period = factor(spatial_period, levels = rev(unique(spatial_period)))) %>% 
  arrange(spatial_period)

mean.bar <- plot_rating(mean.rating.df, "Mean Rating") +
  facet_wrap(~su, ncol = 1)
```

```{r}
mean.rating.exc <- exclude_insufficient(mean.rating.df, spatial_period, method, su)

mean.bar.exc <- plot_rating(mean.rating.exc, "Mean Rating (Insufficient Excluded)") +
  facet_wrap(~su, ncol = 1)
```

```{r, fig.width = 8, fig.height=10}
cowplot::plot_grid(mean.bar, mean.bar.exc, ncol = 1)
```

```{r}
join_cells(bibi.periods, su.df, "huc12", "stream_miles") %>% 
  mean_cell_rating(huc12, stream_miles) %>% 
  spatial_stacked()
```


## 3) Ratings Weighted by Stream Mile


```{r}
stream.weight.rating.df <- loop_cell_rating(bibi.periods, su.df,
                                            unit.col = "stream_miles",
                                            su.vec = c("huc12"),
                                            job = "weighted_rating") %>% 
  arrange(spatial_period) %>% 
  mutate(spatial_period = factor(spatial_period, levels = rev(unique(spatial_period)))) %>% 
  arrange(spatial_period)

sm.bar <- plot_rating(stream.weight.rating.df, "Weighted by Stream Mile") +
  facet_wrap(~su, ncol = 1)
```

```{r}
stream.weight.rating.exc <- exclude_insufficient(stream.weight.rating.df,
                                                 spatial_period, method, su)

sm.bar.exc <- plot_rating(stream.weight.rating.exc,
                          "Weighted by Stream Mile (Insufficient Excluded)") +
  facet_wrap(~su, ncol = 1)
```

```{r, fig.width = 8, fig.height=10}
cowplot::plot_grid(sm.bar, sm.bar.exc , ncol = 1)
```

```{r}
join_cells(bibi.periods, su.df, "huc12", "stream_miles") %>% 
  weighted_cell_rating(huc12, stream_miles) %>% 
  spatial_stacked()
```


## 4) Random Sample

```{r, fig.width=8, fig.height=10}
set.seed(1018) # sample(1:10000, 1)
rand.samp.rating.df <- loop_sample_ratings(bibi.periods, su.df,
                                           su.vec = c("huc12"),
                                           su.density.quantile = 0.1,
                                           reps = 1,
                                           boot.sample.vec = 100,
                                           spatial.vec,
                                           periods.vec) %>% 
  arrange(spatial_period) %>% 
  mutate(spatial_period = factor(spatial_period, levels = rev(unique(spatial_period)))) %>% 
  arrange(spatial_period) %>% 
  filter(bootstrap_sample == 100)

rand.bar <- plot_rating(rand.samp.rating.df,
                        title = "Random Sample",
                        y = percentage) +
  facet_wrap(~su, ncol = 1)
```

```{r}
rand.samp.rating.exc <- exclude_insufficient(rand.samp.rating.df,
                                             spatial_period, su)

rand.bar.exc <- plot_rating(rand.samp.rating.exc,
                            "Random Sample (Insufficient Excluded)") +
  facet_wrap(~su, ncol = 1)
```

```{r, fig.width = 8, fig.height=10}
cowplot::plot_grid(rand.bar, rand.bar.exc , ncol = 1)
```

## 5) Bootstrap
```{r, fig.width=8, fig.height=10}
set.seed(1018) # sample(1:10000, 1)
boot.rating.df <- loop_sample_ratings(bibi.periods, su.df,
                                      su.vec = c("huc12"),
                                      su.density.quantile = 0.1,
                                      reps = 100,
                                      boot.sample.vec = 100) %>% 
  arrange(spatial_period) %>% 
  mutate(spatial_period = factor(spatial_period, levels = rev(unique(spatial_period)))) %>% 
  arrange(spatial_period) %>% 
  filter(bootstrap_sample == 100)


boot.bar <- plot_rating(boot.rating.df,
                        title = "Bootstrap",
                        y = percentage) +
  facet_wrap(~su, ncol = 1)
```

```{r}
boot.rating.exc <- exclude_insufficient(boot.rating.df,
                                        spatial_period, su, job = "bootstrap")

boot.bar.exc <- plot_rating(boot.rating.exc,
                            "Bootstrap (Insufficient Excluded)") +
  facet_wrap(~su, ncol = 1) 
```

```{r, fig.width = 8, fig.height=10}
cowplot::plot_grid(boot.bar, boot.bar.exc , ncol = 1)
```

