---
title: "Untitled"
author: "Zachary M. Smith"
date: "January 9, 2018"
output: html_document
---

```{r echo=FALSE}
knitr::opts_chunk$set(eval=evaluate, cache=cache.me, cache.path=cache.dir)
```

This section provides visuals of sample counts per year, which are helpful in defining potential baseline periods.

In subsequent sections the data will be summarized so that each station has one representative sample per the defined baseline sampling period. To count just the potential number of sampling stations (`station_id`), `dplyr::select()` is used to subset the columns to only represent sampling date (`date`), sample aggregation factor(`subspatial`), and station ID (`station_id`). `dplyr::distinct()` is used to remove any duplicate rows. The `bibi.count` data frame will be used to generate the subsequent plots in this section.
```{r}
bibi.count <- bibi.df %>% 
  select(date, subspatial, station_id) %>% 
  distinct()
```

### Basin

Count the number of sampling events per year. The baseline year, 2008, is colored blue.
```{r, fig.width = 8, fig.height = 2}
count.bar <- bibi.count %>% 
  mutate(year = year(date)) %>% 
  select(year, station_id) %>% 
  distinct() %>% 
  group_by(year) %>% 
  summarize(count = n()) %>% 
  ungroup() %>% 
  mutate(baseline = if_else(year == 2008, TRUE, FALSE)) %>% 
  ggplot(aes(year, count, fill = baseline)) +
  scale_fill_manual(values = c("TRUE" = "#0072B2", "FALSE" = "#999999")) +
  geom_bar(stat = "identity") +
  guides(fill = FALSE) +
  xlab("Year") +
  ylab("Count")

count.bar
```

```{r, include=FALSE}
ggsave("count_bar.png",
       count.bar,
       path = file.path(rprojroot::find_rstudio_root_file(), "markdown/presentation/figures"),
       width = 8,
       height = 2)
```

### Region

Split the count the number of sampling events per year by region. The baseline year, 2008, is colored blue.
```{r, fig.width = 8, fig.height = 2}
region.bar <- bibi.count %>% 
  filter(subspatial %in% c("inland", "coast")) %>% 
  mutate(year = year(date)) %>% 
  select(-date) %>% 
  distinct() %>% 
  group_by(year, subspatial) %>% 
  summarize(count = n()) %>% 
  ungroup() %>% 
  mutate(baseline = if_else(year == 2008, TRUE, FALSE)) %>% 
  arrange(subspatial) %>% 
  ggplot(aes(year, count, fill = baseline)) +
  scale_fill_manual(values = c("TRUE" = "#0072B2", "FALSE" = "#999999")) +
  geom_bar(stat = "identity") +
  guides(fill = FALSE) +
  facet_wrap(~ subspatial, ncol = 2) +
  xlab("Year") +
  ylab("Count")

region.bar
```

```{r, include=FALSE}
ggsave("count_region_bar.png",
       region.bar,
       path = file.path(rprojroot::find_rstudio_root_file(), "markdown/presentation/figures"),
       width = 8,
       height = 2)
```

### Bioregion

Split the count the number of sampling events per year by bioregion. The baseline year, 2008, is colored blue.
```{r, fig.width = 8, fig.height = 8}
bioregion.bar <- bibi.count %>% 
  filter(!subspatial %in% c("inland", "coast")) %>% 
  mutate(year = year(date)) %>% 
  select(-date) %>% 
  distinct() %>% 
  group_by(year, subspatial) %>% 
  summarize(count = n()) %>% 
  ungroup() %>% 
  mutate(baseline = if_else(year == 2008, TRUE, FALSE)) %>% 
  arrange(subspatial) %>% 
  ggplot(aes(year, count, fill = baseline)) +
  scale_fill_manual(values = c("TRUE" = "#0072B2", "FALSE" = "#999999")) +
  geom_bar(stat = "identity") +
  guides(fill = FALSE) +
  facet_wrap(~ subspatial, ncol = 2) +
  xlab("Year") +
  ylab("Count")

bioregion.bar
```

```{r, include=FALSE}
ggsave("count_bioregion_bar.png",
       bioregion.bar,
       path = file.path(rprojroot::find_rstudio_root_file(), "markdown/presentation/figures"),
       width = 8,
       height = 8)
```