---
title: "Untitled"
author: "Zachary M. Smith"
date: "January 8, 2018"
output: html_document
---

### Grid Data

Several grid scales will be evaluated for aggregating Chessie BIBI ratings and ultimately relating ratings to stream miles. The bioregions (`bioregions`) defined by Smith et al. (2017) represent one of the potential grids. The remaining grids are obtained from NHDPlus V1 Hydrologic Unit Codes (HUCs) and catchments. ArcGis was used to clip all of the grids to represent only the areas in the Chesapeake Bay basin.
```{r}
grid.df <- data.table::fread(file.path(project.dir, "data/grid_info/stream_network_grid_master.csv"),
                             colClasses = c(HUC4 = "character",
                                            HUC6 = "character",
                                            HUC8 = "character",
                                            HUC10 = "character",
                                            HUC12 = "character")) %>% 
  clean_up() %>% 
  rename(stream_segment = comid)
```


Join the rating data (`bibi.df`) with the HUC 12 data (`huc12.df`) to obtain the square acres each HUC 12 represents (`acres`) and the states that the HUC 12 falls within (`states`).
```{r}
bibi.df <- left_join(bibi.df, huc12.df, by = c("huc8", "huc10", "huc12"))
```