---
title: "Developing a 2008 Baseline for the CBP Stream Health Outcome Indicator: DRAFT"
author: "Zachary M. Smith"
output:
#  word_document:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '6'
    toc_float: yes
bibliography: citations.bib
link-citations: true
---

`r Sys.Date()`

```{r, echo=FALSE, cache=FALSE}
evaluate <- TRUE
cache.me <- TRUE
cache.dir <- "sections/cache"
run.cedr.acquisition <- FALSE
```

# Abstract
```{r child = 'sections/abstract.Rmd', eval=TRUE}

```

# Goals

* Select a robust baseline period.
* Select a sampling grid representative of stream miles.
* Identify a standard spatial sampling density.

# Background
```{r child = 'sections/background.Rmd', eval=TRUE}

```

# R
```{r child = 'sections/methods_r.Rmd', eval=TRUE}

```

# Data Acquisition
```{r child = 'sections/data_acquisition_cedr.Rmd', eval=TRUE}

```

# Prepare Data
```{r child = 'sections/prepare_data.Rmd', eval=TRUE}

```

# Maps
```{r child = 'sections/map_functions.Rmd', eval=FALSE}

```

```{r child = 'sections/map_import.Rmd', eval=FALSE}

```

```{r child = 'sections/map_plots.Rmd', eval=FALSE}

```

# Bootstrap Sampling
```{r child = 'sections/bootstrap_overview.Rmd', eval=TRUE}

```

## Functions
```{r child = 'sections/bootstrap_functions.Rmd', eval=TRUE}

```

## Grids Explored

The grids explored are organized from largest to smallest cell sizes.

### Bioregion
```{r child = 'sections/bootstrap_bioregion.Rmd', eval=TRUE}

```

### HUC 6
```{r child = 'sections/bootstrap_huc6.Rmd', eval=TRUE}

```

### HUC 8
```{r child = 'sections/bootstrap_huc8.Rmd', eval=TRUE}

```

### HUC 10
```{r child = 'sections/bootstrap_huc10.Rmd', eval=TRUE}

```

### HUC 12
```{r child = 'sections/bootstrap_huc12.Rmd', eval=TRUE}

```

### Catchment
```{r child = 'sections/bootstrap_catchment.Rmd', eval=TRUE}

```

# Citations

