---
title: "Developing a 2008 Baseline for the CBP Stream Health Outcome Indicator: DRAFT"
author: "Zachary M. Smith"
output:
#  word_document:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '6'
    toc_float: yes
bibliography: citations.bib
link-citations: true
---

`r Sys.Date()`

```{r, echo=FALSE, cache=FALSE}
evaluate <- TRUE
cache.me <- TRUE
cache.dir <- file.path(rprojroot::find_rstudio_root_file(), "sections/cache/")
run.cedr.acquisition <- FALSE
```

```{r echo=FALSE}
knitr::opts_chunk$set(cache.path=cache.dir)
```

# Abstract
```{r child = 'sections/prep/abstract.Rmd', eval=TRUE}

```

# Goals

* Select a robust baseline period.
* Select a spatial unit to represent stream miles.
* Identify a standard spatial sampling density.

# Background
```{r child = 'sections/prep/background.Rmd', eval=TRUE}

```

# R
```{r child = 'sections/prep/methods_r.Rmd', eval=TRUE}

```

# Data Acquisition
```{r Data Acquisition, child = 'sections/data_acquistion/data_acquisition_cedr.Rmd', eval=FALSE}

```

# Data Exploration

## Sample Counts
```{r child = 'sections/prep/prepare_sample_counts.Rmd', eval=TRUE}

```

## Maps
```{r child = 'sections/maps/map_functions.Rmd', eval=FALSE}

```

```{r child = 'sections/maps/map_import.Rmd', eval=FALSE}

```

```{r child = 'sections/maps/map_plots.Rmd', eval=FALSE}

```

# Prepare Data
```{r child = 'sections/prep/prepare_data.Rmd', eval=TRUE}

```

# Methods

The R functions below should enable anyone to re-create the results found in this document. Associated with each function is a brief description of what the function does and often some explanation of how to use the function.

```{r child = 'sections/ratings/rating_functions.Rmd', eval=TRUE}

```

## Bootstrap Methods

```{r child = 'sections/ratings/bootstrap/bootstrap_overview.Rmd', eval=TRUE}

```

```{r child = 'sections/ratings/bootstrap/bootstrap_functions.Rmd', eval=TRUE}

```

# Results

```{r child = 'sections/ratings/ratings.Rmd', eval=TRUE}

```

### Spatial Units Explored

The spatial units explored are organized from largest to smallest cell sizes.

#### HUC 8
```{r child = 'sections/ratings/bootstrap/bootstrap_huc8.Rmd', eval=TRUE}

```

#### HUC 10
```{r child = 'sections/ratings/bootstrap/bootstrap_huc10.Rmd', eval=TRUE}

```

#### HUC 12
```{r child = 'sections/ratings/bootstrap/bootstrap_huc12.Rmd', eval=TRUE}

```

#### Catchment
```{r child = 'sections/ratings/bootstrap/bootstrap_catchment.Rmd', eval=TRUE}

```

## Random Forest


# Discussion
```{r child = 'sections/discussion/discussion.Rmd', eval=TRUE}

```

# Conclusions and Recommendations


# Citations

