---
title: "Untitled"
author: "Zachary M. Smith"
date: "March 13, 2018"
output: ioslides_presentation
---

## Random Forest {.smaller}

__Process__

1. Chessie BIBI ratings are aggregated by catchment
    + 1:24,000 high resolution [SHEDS](www.ecosheds.org)
2. A random forest model was developed to predict Chessie BIBI ratings based on the geospatial predictor variables (e.g., land cover, soil, precipitation, and number of dams)
3. Using the model predict across the entire Chesapeake Bay basin if an area represents FairGood, Poor, or Uncertain

## Random Forest

```{r, fig.width=6, fig.height=6, fig.show='hold', fig.align='center'}
imager::load.image(file.path(project.dir, "markdown/presentation/figures/results/random_forest/rf_prediction_uncertain.jpg")) %>% 
  plot(axes = F)
```

## Random Forest

__Discussion__

## Random Forest

```{r, fig.width=6, fig.height=6, fig.show='hold', fig.align='center'}
imager::load.image(file.path(project.dir, "markdown/presentation/figures/results/random_forest/rf_prediction_uncertain.jpg")) %>% 
  plot(axes = F)
```