---
title: "Calculating and Mapping Stream Biological Health"
author: "Zachary M. Smith"
date: "April 5, 2018"
output: ioslides_presentation
css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
project.dir <- rprojroot::find_rstudio_root_file()
```

## Issues to Keep in Mind {.smaller}

* How to equate a Chessie BIBI rating to stream miles
    + One sampling event equals X amount of miles?
    + Aggregate by a spatial unit (e.g., HUC 12) and use sampling events within a spatial unit cell to represent all stream miles within that spatial unit cell?
* Spatial Influence
    + Sampling density and frequency (Sampling Rounds)
    + The amalgamation of two or more random sampling designs does __NOT__ equate to a single random sampling design
    + Stream access, generally, leads to a larger proportion of streams sampled in urban areas
* Temporal Influence
    + Sampling density and frequency (Sampling Rounds)
    + 2000-2008 vs. 2004-2008 baseline periods (Other Suggestions?)
* Index Resolution
    + Region indices vs. Bioregion indices
    
## Six Methods

1. No Spatial Aggregation
2. Watershed Mean Score
3. Ratings Weighted by Stream Mile
4. Random Sample
5. Bootstrap
6. Random Forest


```{r child = 'sections/no_spatial.Rmd', eval=TRUE}

```

```{r child = 'sections/watershed_mean_score.Rmd', eval=TRUE}

```

```{r child = 'sections/weighted_rating.Rmd', eval=TRUE}

```

```{r child = 'sections/random_sample.Rmd', eval=TRUE}

```

```{r child = 'sections/bootstrap.Rmd', eval=TRUE}

```

```{r child = 'sections/random_forest.Rmd', eval=TRUE}

```

---- {.flexbox .vcenter}


__Break for Lunch__

__12:00-1:00__


