---
title: "Untitled"
author: "Zachary M. Smith"
date: "December 27, 2017"
output: html_document
---

```{r echo=FALSE}
knitr::opts_chunk$set(eval=evaluate, cache=cache.me)
```

## Import Shapefiles

Import shapefiles as spatial polygon data frames for visualizing different grids that could be used during the analysis.
```{r}
basin.poly <- readOGR(file.path(project.dir, "data/gis/NHDPlus/shapefiles/basin.shp"),
                      verbose = FALSE)

bioregion.poly <- readOGR(file.path(project.dir, "data/gis/NHDPlus/shapefiles/bioregion.shp"),
                      verbose = FALSE) 

huc6.poly <- readOGR(file.path(project.dir, "data/gis/NHDPlus/shapefiles/huc6.shp"),
                      verbose = FALSE)

huc8.poly <- readOGR(file.path(project.dir, "data/gis/NHDPlus/shapefiles/huc8.shp"),
                      verbose = FALSE)

huc10.poly <- readOGR(file.path(project.dir, "data/gis/NHDPlus/shapefiles/huc10.shp"),
                      verbose = FALSE)

huc12.poly <- readOGR(file.path(project.dir, "data/gis/NHDPlus/shapefiles/huc12.shp"),
                      verbose = FALSE)

catchment.poly <- readOGR(file.path(project.dir, "data/gis/NHDPlus/shapefiles/catchment.shp"),
                      verbose = FALSE)
```

State polygons can be imported using the function `mapps::map_data()`. `dplyr::unite()` is used to concatenate the `region` and `subregion` columns into a grouped region column (`gregion`). The grouped region column (`greion`) is necessary to maintain polygon integrity because without this column the polygons will often be distorted when plotting.
```{r}
states.vec <- c('Maryland', "Virginia", "West Virginia",
                "Delaware", "Pennsylvania", "New York",
                "District of Columbia")

state.df <- map_data('state', region = states.vec) %>% 
  unite(gregion, region, subregion, remove = FALSE)
```

Convert the state data frame (`state.df`) to a spatial polygon data frame using `prep_spdf()` (see [Map Functions]), so that it can be clipped by other polygons in subsequent code chunks.
```{r}
state.spdf <- prep_spdf(state.df, gregion)
```

Clip, using `rgeos::gIntersection()`, the state polygons (`state.spdf`), so they are only represented as areas that fall within the Chesapeake Bay basin (`basin.poly`). Use `sp::SpatialPolygonsDataFrame()` to convert the intersected polygons to spatial polygon data frames.
```{r}
state.clip <- gIntersection(state.spdf, basin.poly, byid = TRUE) %>% 
  SpatialPolygonsDataFrame(data.frame(id = names(.), 
                                           row.names = names(.)))
```

